Reproductive Allocation Calculations

### 2015-03-24 leaflifespan.Rmd
Move functions into utils file, move packages into remake.yml

```{r, echo=FALSE,results='hide', message=FALSE}
# load packages
library(tidyr)
library(dplyr)
#library(lme4)

#load files

LMA <- read.csv("data/LMA_short.csv",header=TRUE,as.is=TRUE,stringsAsFactors=FALSE)%>%
  filter(species!="" & species!=" ")
#InvestmentSummary <- remake::make("ReproductionAllocation_all")
#names(InvestmentSummary)[1] <- "tag_ID"

#define functions
se <- function(x) sd(x)/sqrt(length(x))
con95 <- function(x) (sd(x)/sqrt(length(x)))*1.96

# leaf lifespan data
raw_leaf_lifespan <- read.csv("data/leaderGrowthLL.csv",header=TRUE,as.is=TRUE,stringsAsFactors=FALSE) %>%
  filter(dont_use!="dead" & dont_use!="dont_use")

@@

# get average leaves per length for spp
leaves_per_length_spp <- read.csv("data/leaves_per_length.csv", header=TRUE, as.is=TRUE,stringsAsFactors=FALSE) %>%
  filter(use=="yes") %>%
  mutate(count_per_length = leaf_count/length_mm) %>%
  group_by(species) %>%
  summarise(count_per_length = mean(count_per_length, na.rm=TRUE))

@@

write.csv(raw_leaf_lifespan, file = "raw_leaf_lifespan.csv")
```

@@

#creating summary files
#leaf lifespan summary by species, age
LLSummary <- raw_leaf_lifespan %>%
  filter(LL_bd>0 & LL_bd<6 & LL_birth<8 & LL_death<8) %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth,new_length, lvs_end)

@@

#leaf lifespan summary by species

LL_spp <- raw_leaf_lifespan %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth)

@@

#merge LMA data into LLSummary dataframe
LLSummary <- merge(LLSummary, LMA_summary, by.x=c("species","age"), by.y=c("species","age"))
LL_spp <- merge(LL_spp, LMA_spp, by.x=c("species"), by.y=c("species"))

#plotting and stats
#creating lists of information for plotting symbols, colors, labels
pch.spp <- c(0,1,2,15,16,17,18,0,1,2,15,16,17,18)
col.spp <- c("red", "dark green", "blue", "purple", "orange", "light green","salmon","turquoise","grey", "black","light blue","brown","yellow","salmon")
col.age <-c("purple","pink","green","orange","blue","light blue")
pch.age <- c(1,1,16,16,16,16)
labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU") 
labels.age <- c("1.4","2.4","5","7","9","32")

@@

labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")

@@

#models with leaf counts per age
The number of leaves on a leader is slightly affected by age however, with a decrease in leaf count along the leader with increasing age. This is what gives the perception of fewer leaves in some older plants.

@@

labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")

plot_lfcountvAge <- function(spp) {
  plot(lvs_end ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leaf count along shoot", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp]) 
  points(lvs_end_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((lvs_end_mean+lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((lvs_end_mean-lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

@@

#models with shoot extenstion by age
These models show that for some species, shoot extension decreases enormously in the older plants. This effect is larger than with number of leaves, indicating that some of these individuals have more leaves per length as they get older

```{r, echo=FALSE,results='hide', message=FALSE}
labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")

@@

#plots regressing variables against one another

>>>>

```{r, echo=FALSE,results='hide', message=FALSE}
# load packages

library(remake)
create_bindings()

# InvestmentSummary <- InvestmentSummary
# names(InvestmentSummary)[1] <- "tag_ID"

LMA <- read.csv("data/LMA_short.csv",header=TRUE,stringsAsFactors=FALSE)%>%
  filter(species!="" & species!=" ")

wood <- read.csv("data/woodDensity.csv", header=TRUE,stringsAsFactors=FALSE)%>%
  filter(use=="use")

seedsize <- read.csv("data/seedsize.csv", header=TRUE,stringsAsFactors=FALSE)

# leaf lifespan data
raw_leaf_lifespan <- read.csv("data/leaderGrowthLL.csv",header=TRUE,stringsAsFactors=FALSE) %>%
  filter(dont_use!="dead" & dont_use!="dont_use")

@@

# get average leaves per length for spp
leaves_per_length_spp <- read.csv("data/leaves_per_length.csv", header=TRUE, stringsAsFactors=FALSE) %>%
  filter(use=="yes") %>%
  mutate(count_per_length = leaf_count/length_mm) %>%
  group_by(species) %>%
  summarise(count_per_length = mean(count_per_length, na.rm=TRUE))

@@

#write.csv(raw_leaf_lifespan, file = "raw_leaf_lifespan.csv")
```

@@

#creating summary files
#leaf lifespan summary by species, age
LLSummary <- raw_leaf_lifespan %>%
  filter(LL_bd>0 & LL_bd<6 & LL_birth<8 & LL_death<8) %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth,new_length, lvs_end, lvs_end_total)

@@

#wood density summary by species
Wood_summary <- wood %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), density)
names(Wood_summary)<-c("species","WD_mean","WD_se","WD_length")

#leaf lifespan summary by species
LL_spp <- raw_leaf_lifespan %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth)

@@

#merge LMA, wood density, seed size data into LLSummary dataframe
LLSummary <- merge(LLSummary, LMA_summary, by.x=c("species","age"), by.y=c("species","age"))
LLSummary <- merge(LLSummary, Wood_summary, by.x=c("species"), by.y=c("species"))
LLSummary <- merge(LLSummary, seedsize, by.x=c("species"), by.y=c("species"))
LL_spp <- merge(LL_spp, LMA_spp, by.x=c("species"), by.y=c("species"))
LL_spp <- merge(LL_spp, Wood_summary, by.x=c("species"), by.y=c("species"))
LL_spp <- merge(LL_spp, seedsize, by.x=c("species"), by.y=c("species"))
#still to add, max H from Investment data, absolute and relative growth measures

#plotting and stats
#creating lists of information for plotting symbols, colors, labels
pch.spp <- c(0,1,2,15,16,17,18,0,1,2,15,16,17,18)
col.spp <- c("red", "dark green", "blue", "purple", "orange", "light green","salmon","turquoise","grey", "black","light blue","brown","yellow","salmon")
col.age <-c("purple","pink","green","orange","blue","dark green")
pch.age <- c(1,1,16,16,16,16)
labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU") 
labels.age <- c("1.4","2.4","5","7","9","32")

@@

@@

#models with leaf counts per age
The number of leaves on a leader is slightly affected by age however, with a decrease in leaf count along the leader with increasing age. This is what gives the perception of fewer leaves in some older plants. I'm not sure why this pattern looks stronger than leaf lifespan.

@@

plot_lfcountvAge <- function(spp) {
  plot(lvs_end_total ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leaf count along shoot", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp]) 
  points(lvs_end_total_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((lvs_end_total_mean+lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((lvs_end_total_mean-lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

@@

#models with shoot extenstion by age
These models show that for some species, shoot extension decreases enormously in the older plants. This effect is larger than with number of leaves.

@@

#plots regressing variables against one another
```{r, echo=FALSE,results='hide', message=FALSE}
plot(change_shoot_area~new_length,data=subset(raw_leaf_lifespan,change_shoot_area<1000000), log="xy",col=col.age[as.factor(age)], pch=16, cex=.8, xlab="shoot extension (mm)", ylab="increased shoot cross sectional area")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(change_shoot_area~new_length +age,data=subset(raw_leaf_lifespan,change_shoot_area<1000000))
summary(mod)

raw_leaf_lifespan$lvs_end_total <- raw_leaf_lifespan$lvs_end + raw_leaf_lifespan$lvs_new
plot(lvs_end_total~lvs_start,raw_leaf_lifespan, log="xy",col=col.age[as.factor(age)], pch=16, cex=.8, xlab="starting leaf count", ylab="ending leaf count")

legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(lvs_start~lvs_end_total+age +age*lvs_end_total, raw_leaf_lifespan)
summary(mod)
#IS THERE AN EASY WAY TO ADD A 1:1 LINE TO THIS PLOT?

plot(LL_birth_mean~LMA_mean,data=subset(LLSummary,species!="HATE"),col=col.age[as.factor(age)], pch=16, cex=.8, log="xy",xlab="LMA", ylab="leaf lifespan (calculated from leaf births")
points(LL_birth_mean~LMA_mean,data=subset(LLSummary,species=="HATE"),col="red", pch=19, cex=.8)
